A Study on Neural Models for Target-Based Computer-Assisted Musical Orchestration

### 2021-04-08 csmc_mume_2020/paper/JCMS_2021/paper.tex
More edits to text mostly

This paper is an extended version of a paper by the same title in \textit{The 2020 Joint Conference on AI Music Creativity} \cite{Cella2020c}. This version includes additional figures and extended information on our baseline testing and orchestration process. The codebase for this paper can be found at \url{https://github.com/dzluke/DeepOrchestration} and you can listen to audio examples at \url{https://dzluke.github.io/DeepOrchestration/}.

@@

In other words, the classifiers learn how to take a complex combination of pitches and timbres and deconstruct it into its original parts.

In order to have a baseline to compare our classification results and to get a sense on the complexity of the problem, we used various parametric classifiers along with the neural models.

As described in Sec. \ref{sec:baseline}, three baseline classifiers have been trained on simplified versions of our problem. They have been trained to either identifying only the instruments present, or the instruments and pitch classes present.

A complete orchestration solution, however, would normally be made by triples of instrument-pitch-dynamics. It is difficult to frame this problem as classification, since we would need to have a very high number of classes. Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in \ref{sec:dataset}), each class would be represented by a single sample. For these reasons, our models have been trained on simplified versions of the original problem and we used different strategies to provide the remaining information needed to generate orchestrations.

Two neural classifiers have been trained on the more complex problem of recognizing instrument-pitch pairs, for a total of 424 classes (i.e. 424 unique instrument-pitch pairs). We want to point out, however, that this is still a simplified problem compared to assisted orchestration. Even in the case of instrument-pitch pair classification, indeed, we would miss the dynamics of the note. To cope with this problem, in the orchestration phase, we quantize the probability functions generated by the models. When a network is given a target sound as input, it will attempt to apply the same classification rules, outputting a vector of the probabilities of each sample being in the target sound. Since the classes only encode instrument-pitch pairs and since we also need the dynamics, we quantize the probability function generated by the network into three echelons coresponding, respectively, to the dynamic levels \emph{pianissimo}, \emph{mezzoforte} and \emph{fortissimo} (see section \ref{sec:orchestration}). By taking the instruments and pitches that have the highest probability and by using quantized probabilities to retrieve dynamics, a full orchestrated solution can be created.  Figure \ref{fig:method} illustrates the described ideas.

@@

The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit, whose structure is inspired by the success in \cite{Salamon17}. CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey17}. LSTM units provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.

>>>>

This paper is an extended version of a paper by the same title in \textit{The 2020 Joint Conference on AI Music Creativity} \cite{Cella2020c}. This version includes additional figures and extended information on our baseline testing and orchestration process, among other additions. The codebase for this paper can be found at \url{https://github.com/dzluke/DeepOrchestration} and you can listen to audio examples at \url{https://dzluke.github.io/DeepOrchestration/}.

@@

In order to have a baseline to compare our classification results and to get a sense of the complexity of the problem, we used various parametric classifiers along with the neural models. As described in section \ref{sec:baseline}, three baseline parametric classifiers have been trained on simplified versions of our problem. They have been trained to either identify only the instruments present, or the instruments and pitch classes present, two simplified versions of our problem.

A complete orchestration solution, however, would normally be made by triples of instrument-pitch-dynamics. It is difficult to frame this problem as classification, since we would need to have a very high number of classes. Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in section \ref{sec:dataset}), each class would be represented by a single sample. For these reasons, our models have been trained on simplified versions of the original problem and we used different strategies to provide the remaining information needed to generate orchestrations.

Two neural classifiers have been trained on the problem of recognizing instrument-pitch pairs, which has a total of 424 classes (i.e. 424 unique instrument-pitch pairs). We want to point out, however, that this is still a simplified problem compared to assisted orchestration. Even in the case of instrument-pitch pair classification, indeed, we would miss the dynamics of the note. To cope with this problem, in the orchestration phase, we quantize the probability functions generated by the models. When a network is given a target sound as input, it will attempt to apply the same classification rules, outputting a vector of the probabilities of each sample being in the target sound. Since the classes only encode instrument-pitch pairs and since we also need the dynamics, we quantize the probability function generated by the network into three echelons coresponding, respectively, to the dynamic levels \emph{pianissimo}, \emph{mezzoforte} and \emph{fortissimo} (see section \ref{sec:orchestration}). By taking the instruments and pitches that have the highest probability and by using quantized probabilities to retrieve dynamics, a full orchestrated solution can be created.  Figure \ref{fig:method} illustrates the described ideas.

@@

The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit, whose structure is inspired by the success in~\cite{Salamon17}. CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey17}. LSTM units provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.
